---
title: "Report of Previous Study"
output:
  html_document:
    toc: true
    toc_float: yes
---


## Description Table

Table 1 Distribution of epidemiology and clinical data of all mpox cases between 2023-2024

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(forcats)
library(tidyverse)

#table1
setwd("C:/Users/dhihr/OneDrive/mpox")
metadata <- read.csv('dat_indo.csv')
metadata <- metadata %>%
  mutate(age_cat = case_when(
    Age <= 12 ~ "0-12",
    Age <= 19 ~ "13-19",
    Age <= 64 ~ "20-64",
    Age > 64 ~ "65+",
    TRUE ~ "Missing",
  ))
metadata <- rename(metadata, label = Virus.name)
#metadata$Submitting.lab <- 'Balai Besar Laboratorium Biologi Kesehatan'
#metadata2 <- read.csv('metadata_samp.csv')
#metadata2 <- metadata2 %>% select(label, location )
#metadata <- metadata %>% select(-location)
#metadata <- left_join(metadata, metadata2, by = "label")
metadata$area <- ifelse(metadata$area == 'Jabodetabek', 'The Greater Jakarta', metadata$area)

library(table1)

table <- metadata
#table(table$Submitting.lab)
#table$Sex[table$Sex == "C"] <- "Missing"
#table[table$Submitting.lab == "",] <- "Missing"
#table[table$Clade == "",] <- "Non-Detected"
table$Sex <- ifelse(table$Sex == "M", "Male", 
                    ifelse(table$Sex == "F", "Female", table$Sex))

#conditional for separator commas

render.categorical <- function(x, ...) {
  c("", sapply(stats.apply.rounding(stats.default(x)), function(y) with(y,
                                                                        sprintf("%s (%s%%)", prettyNum(FREQ, big.mark=","), PCT))))
}

render.strat <- function (label, n, ...) {
  sprintf("<span class='stratlabel'>%s<br><span class='stratn'>(N=%s)</span></span>", 
          label, prettyNum(n, big.mark=","))
}


# Reorder levels for 'Clade'
table$Clade <- table$Clade %>%
  fct_infreq() %>%
  fct_relevel("Non-Detected", after = Inf)

# Reorder factor levels for multiple columns
table <- table %>%
  mutate(across(c(Sex, age_cat, wgs, Severity, Sex.Orientation, Status, HIV, Travel_hist, loc, area, Isolation), ~ fct_infreq(.)))


#labelling
label(table$Sex) <- "Sex"
label(table$loc) <- "Origin Report"
label(table$wgs) <- "Status WGS"
label(table$Severity) <- "Severity Criteria"
label(table$Sex.Orientation) <- "Sex Orientation"
label(table$area) <- "Location"
label(table$Clade) <- "Clade"
label(table$Status) <- "Final Status"
label(table$Travel_hist) <- "Travel History"
label(table$HIV) <- "HIV Status"
label(table$age_cat) <- "Age Group"
label(table$los) <- "Length of Treatment (days)"
label(table$Isolation) <- "Isolation Status"

#table
tab1 <- table1(~Sex + age_cat + Sex.Orientation +Travel_hist + loc + area +
                 Severity + Clade + HIV + Isolation + los + Status| wgs
               , data=table, render.categorical=render.categorical, 
               overall=c(left="Total"), render.strat=render.strat, 
               caption = ' ')
tab1
```

## Description Table by Clade

Table 2 Descriptive summary of the relationships between different Clades (IIb sub-clade or lineages) and clinical conditions, 

```{r, echo = FALSE, warning = FALSE, message = FALSE}
#table group by clade
table2 <- table %>% filter(Clade != "Non-Detected")
table2$Severity <- relevel(table2$Severity, ref = "Mild")
table2$los_cat <- as.factor(ifelse(table2$los >= 14, "≥ 14 days", "≤ 14 days"))
table2$los_cat <- relevel(table2$los_cat, ref = "≤ 14 days")
table2$HIV <- relevel(table2$HIV, ref = "No")
table2$ART_treatment <- as.factor(as.character(ifelse
                                               (table2$HIV == "No", "Non HIV", table2$ART_treatment)))
table2$ART_treatment <- table2$ART_treatment %>%
  fct_infreq() %>%
  fct_relevel("Non HIV", after = Inf)
table2$ART_treatment <- relevel(table2$ART_treatment, ref = "Non ART")

label(table2$los_cat) <- "Length of Treatment"
tab3 <- table1(~Severity + los_cat + HIV + ART_treatment | Clade
               , data=table2, render.categorical=render.categorical, 
               overall=c(left="Total"), render.strat=render.strat, 
               caption = ' ')
tab3
```

## Epidemic Curve

Figure 1 The figure A. shows the epi-curve of mpox cases monthly in Indonesia, the cases were grouped into non-WGS and WGS samples. Meanwhile, the B. illustrates the trend of sub-clade among the WGS cases 

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Convert the date from character to Date type if necessary
metadata$Date.Onset <- as.Date(metadata$Date.Onset , format = "%m/%d/%Y")
metadata$date_diagnosis <- as.Date(metadata$date_diagnosis , format = "%m/%d/%Y")
metadata$date_discharge <- as.Date(metadata$date_discharge , format = "%m/%d/%Y")
metadata$Date.Onset <- as.Date(ifelse(is.na(metadata$Date.Onset), metadata$date_diagnosis, metadata$Date.Onset))

#Cases
summary_table <- metadata %>%
  group_by(wgs,Date.Onset) %>%
  summarise(count = n())

# Convert dates to the first date of each month
summary_table <- summary_table %>%
  mutate(first_day_of_month = floor_date(Date.Onset, "month"))
time <- summary_table %>%
  group_by(wgs, first_day_of_month) %>%
  summarize(
    Cases = sum(count))


epicurve <- ggplot(time, aes(x = first_day_of_month, y = Cases, fill = wgs)) +
  geom_bar(stat = "identity", alpha = 0.8) +
  labs(
    title = " ",
    x = "Date",
    y = "Number of Cases",
    fill = "Category Cases"
  ) + ggtitle("A. Monthly Cases") +scale_x_date(
    date_breaks = "3 month",  # Set the interval for breaks
    date_labels = "%b-%Y",  # Customize the format (e.g., Aug 22, 2022)
    limits = as.Date(c('2023-01-01','2024-08-01'))
  ) + theme_classic() + theme(
    legend.position = "top")
epicurve <- epicurve + theme(panel.grid.major = element_blank(),  # Removes major grid lines
                 panel.grid.minor = element_blank(),  # Removes minor grid lines
                 axis.title.x=element_blank(),
                 axis.text.x=element_blank(),
                 axis.ticks.x=element_blank())

#Clade
summary_table <- metadata %>%
  group_by(Clade,Date.Onset) %>%
  summarise(count = n())

# Convert dates to the first date of each month
summary_table <- summary_table %>%
  mutate(first_day_of_month = floor_date(Date.Onset, "month"))
time <- summary_table %>%
  group_by(Clade, first_day_of_month) %>%
  summarize(
    Cases = sum(count))
time <- filter(time, Clade != "Non-Detected")

clade <- ggplot(time, aes(x = first_day_of_month, y = Cases, fill = Clade)) +
  geom_bar(position="fill", stat="identity", alpha = 0.8) +
  labs(
    title = " ",
    x = "Month",
    y = "Proportion of Clades",
    fill = "Clade"
  ) + scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  ggtitle("B. Proportion of Clades") + scale_x_date(
    date_breaks = "3 month",  # Set the interval for breaks
    date_labels = "%b-%Y",  # Customize the format (e.g., Aug 22, 2022)
    limits = as.Date(c('2023-01-01','2024-08-01'))
  ) + theme_classic() + theme(
    legend.position = "top")

library(patchwork)
epicurve / clade
```

## Phylogenetic Tree

Figure 2 The A, displays a phylogenetic tree of cases from Indonesia, with the x-axis representing the substitution rate with epidemiological data. In contrast, the B. figure illustrates a time-calibrated phylogenetic tree of sequences from ASEAN countries, indicating that  the most recent common ancestor (TMRCA) dates to approximately July 2018. The B tree diverged into two groups or branches: Group 1 (yellow shade) and Group 2 (gray shade). Also, The B tree tips are also divided into clade subgroups or lineages. The C. is the ASEAN Subset where the green label and red line were positioned at the time of introduction estimation (December 2021 and January 2023). Both ASEAN and Indonesia’s tree were rooted against reference (NC_063383.1 ). 

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(ape)
library(ggtree)
library(forcats)
library(tidytree)
library(treeio)
library(tidyverse)
setwd("C:/Users/dhihr/OneDrive/mpox/revisi/tree_time/mpox_timetree")
nexus_tree <- read.nexus("timetree.nexus")


#reroot
#lable_reroot <- 'NC_063383.1'
#tree <- root(nexus_tree, outgroup = lable_reroot, resolve.root = TRUE)
#plot(tree)

metadata<-read.csv('metadata_asean.csv')
metadata$date <- as.Date(metadata$date, format = "%m/%d/%Y")
metadata$id <- sub("^[^/]+/[^/]+/([^/]+).*$", "\\1", metadata$label)
#metadata$id <- seq_len(nrow(metadata))
#metadata$id <- paste0('C',metadata$id)
metadata[metadata$location == 'African', 'location'] <- 'Ref'
# Reorder levels for 'location'
metadata$location <- as.factor(metadata$location)
metadata$location <- metadata$location %>%
  fct_infreq() %>%
  fct_relevel("Ref", after = Inf)

#check nodde
#ggtree(nexus_tree) + geom_text2(aes(label=node), vjust=-0.5, hjust=0.5, size=3, color="blue")

most_recent_date <- max(lubridate::as_date(metadata$date))
#ggtree(nexus_tree, mrsd=most_recent_date) + geom_tiplab() + theme_tree2()
t <- ggtree(nexus_tree, mrsd=most_recent_date)+ theme_tree2()
#tree <- drop.tip(nexus_tree, 'NC_063383.1')
a <- t$data

#by clade
t <- t + geom_hilight(node=87, fill="gold", alpha=0.15) + geom_hilight(node=81, fill="gray", alpha=0.3)
#t %<+% metadata + geom_tippoint(aes(color = clade), size = 2)  +
#  geom_tiplab(aes(label=id), size = 2.2)+  labs(color = "Clade") +
#  theme(legend.position = 'right') 
#by country
asean <- t %<+% metadata + geom_tippoint(aes(color = location), size = 2.2) +  labs(title="B. ASEAN Time-Inference Tree", color = "Country") +
  theme(legend.position = 'right')
#add heatmap
metadata$clade <- as.factor(metadata$clade)
mt <- metadata %>% select(label, clade)
row.names(mt) <- mt$label
mt <- mt %>% select(clade)
#asean <- ggplottree(nexus_tree, metadata, cols=c('location','clade'),
#           cmaps=c('Set1','Set2'), tipsize=3, offset=.3 ,layout='rect')
asean <- gheatmap(asean, mt, offset=0.2, width=0.1, 
         colnames=FALSE, legend_title="Clade") +
  scale_x_ggtree() + 
  scale_y_continuous(expand=c(0, 0.05))

#Indonesia
#reroot
#setwd("C:/Users/dhihr/OneDrive/mpox/revisi/tempest")
#tree <- read.tree("indo_tree_tempest.nwk")
#lable_reroot <- 'hMpxV/Indonesia/JK-NIHRD-MP0024/2022'
#tree <- root(tree, outgroup = lable_reroot, resolve.root = TRUE)
#most_recent_date <- max(lubridate::as_date(ina$date))
#tree <- drop.tip(tree, drop$label)
#t <- ggtree(tree, mrsd=most_recent_date) + theme_tree2()
#t %<+% metadata + geom_tippoint(aes(color = clade), size = 2) + 
  #geom_tiplab(aes(label=id), size = 2)+  labs(color = "Clade") +
  #theme(legend.position = 'right')

#non time tree Indonesia
setwd("C:/Users/dhihr/OneDrive/mpox/revisi/tempest")
tree <- read.tree("indo_tree_tempest.nwk")
t <- ggtree(tree) + theme_tree2() + labs(caption="Substitution rate")

#check nodes
metadata2<-read.csv('metadata_indo.csv')
metadata2$date <- as.Date(metadata2$date, format = "%m/%d/%Y")
metadata2$id <- sub("^[^/]+/[^/]+/([^/]+).*$", "\\1", metadata2$label)
#t %<+% metadata2 + geom_tippoint(aes(color = clade), size = 2) + labs(color = "clade") 
  #theme(legend.position = c(0.1, 0.7)) + geom_treescale(fontsize=6, linesize=2)
  
t <- ggtree(tree) + 
    theme_tree2() + 
    labs(caption = "Substitution rate") +
    labs(color = "clade") +
    geom_treescale(x=0.00005, y=20, fontsize = 4, linesize = 1.5) +
    theme(legend.position = c(0.1, 0.7))
indo <- t %<+% metadata2 + geom_tippoint(aes(color = clade), size = 2.5)  + 
  labs(title = 'A. Indonesia Phylogenetic Tree', color = "Clade") +
    theme(legend.position = 'right')
#add heatmap
metadata2 <- metadata2 %>%
  mutate(across(c(Sex, HIV, Travel, Transmission, Severity), ~ifelse(. == "", " ", .)))
metadata2$HIV <- as.factor(metadata2$HIV)  
metadata2$Severity <- as.factor(metadata2$Severity) 
metadata2$Travel <- as.factor(metadata2$Travel) 
mt2 <- metadata2 %>% select(label, clade, HIV, Severity, Travel)
row.names(mt2) <- mt2$label
mt3 <- mt2 %>% select(Severity, HIV)
mt4 <- mt2 %>% select(HIV)
mt3$Severity <- factor(mt3$Severity, levels = c("Asymptomatic", "Mild", "Moderate", "Severe", "Ref"))
mt4$HIV <- factor(mt4$HIV, levels = c("Yes", "No", "Ref"))
indo1 <- gheatmap(indo, mt3, offset=0.000002, width=0.15,  font.size=0.05,
                  colnames=FALSE, legend_title="Genotype") + 
  coord_cartesian(clip = "off") +
  scale_x_ggtree() + 
  scale_y_continuous(expand=c(0, 0.5)) + scale_fill_manual(
    values = c("Asymptomatic" = "plum2", "Mild" = "hotpink2", "Severe" = "hotpink4" , "Yes" = 'gold4', 'No' = 'gold1', 'Ref' = 'gray95'),
    limits = c('Asymptomatic', 'Mild', 'Severe', 'Yes', 'No', 'Ref'), name = "Categories")
#indo1
#indo2 <- indo1 + ggnewscale::new_scale_fill()
#indo3 <- gheatmap(indo2, mt4, offset=0.000030, width=0.15,  font.size=0.05, colnames_position = "bottom",
#                  colnames=FALSE) + scale_y_continuous(expand=c(0, 0.5)) + scale_fill_viridis_d(option = "C", name = "HIV") 
#indo3

#combine
library(patchwork)

#subset tree
subset <- tree_subset(nexus_tree, node=88, levels_back=0)
#ggtree(subset) + geom_text2(aes(label=node), vjust=-0.5, hjust=0.5, size=3, color="blue")
most_recent_date <- max(lubridate::as_date(metadata$date))
t2 <- ggtree(subset, mrsd=most_recent_date)+ theme_tree2()
a <- t2$data
t2 <- t2 %<+% metadata + geom_tippoint(aes(color = location), size = 3)  +  labs(color = "Country") +
  theme(legend.position = 'right') 
a$text <- ifelse(a$node == 71, "Introduction 1\n2021-12-17", 
                 ifelse(a$node == 87, "Introduction 2\n2022-09-07", NA))

sub <- t2 + geom_vline(xintercept=2021.959, linetype=2, size = 1.1, color = 'firebrick') + 
  geom_vline(xintercept=2023.055, linetype=2, size = 1.1, color = 'firebrick')+
  geom_label(data=a, aes(x=2021.700, y=50, label="Introduction 1\n2021-12-17"),fill = 'darkgreen', size = 3, color = 'white', hjust=0, vjust=0) +
  geom_label(data=a, aes(x=2022.900, y=50, label="Introduction 2\n2023-01-21"),fill = 'darkgreen', size = 3, color = 'white', hjust=0, vjust=0) +
  labs(title = 'C. ASEAN Subset Time-Inference Tree')

indo1
asean
sub

```


